NMRlipids Databank: Overlay Databank of Lipid Membrane Simulations Arising from Open Collaboration

### 2022-08-25 main.tex
Updates from Overleaf

\subsection{NMRlipids databank reveals correlations between membrane properties}
The power of NMRlipids databank to find correlation between membrane properties is exemplified in Fig.~\ref{??} showing how water diffusion through and along the membrane correlate with different properties of systems. While NMRlipids databank enables similar analyses for any membrane properties, these are selected due to their relevance in the development of MRI imaging methods~\cite{??}.

%\subsection{Spin relaxation rates of confined water close to membranes}

%Usefulness of the databank beyond MD simulation experts is demonstrated by analysing water spin relaxation times close to bilayers which are used in MRI imaging.

%Example text under a subsection. Bulleted lists may be used where appropriate, e.g.

%\begin{itemize}
%\item First item
%\item Second item
%\end{itemize}

%\subsubsection*{Third-level section}

%Topical subheadings are allowed.

>>>>

\subsection{Water diffusion anisotropy in membrane systems}
The anisotropic water diffusion through and along of membranes is utilized in MRI imaging \cite{??} and plays a role in the drug translocation through biological material, particularly in skin \cite{hansen13,wen18,nitsche19,roberts21,??}. Dependencies between membrane properties, water permeablity, and anisotropic water diffusion have been recognized \cite{??} but consensus on how water permeability depends on membrane properties, such as area per lipid and thickness, or molecular composition, has not yet been reached despite of several available models~\cite{nitsche13,nitsche16,shinoda16,venable19,??}. MD simulations can be used to elucidate such dependencies, but systematic collection of sufficient amount of data has been challenging because few water permeation events are typically observed in a single MD simulation trajectory \cite{??}. Nevertheless, averaging over large amounts of data available in the NMRlipids databank enables to study the trends of water permeation upon changing memebrane properties.

%averaging Here we demonstrate the power of the NMRlipids databank by analysing how 
Water permeabilities calculated from all the simulations in the NMRlipids databank as a function of area per lipids, thickness, and acyl chain order are shown in  Fig.~\ref{fig:permeability} (a-c). To illustrate how permeability depends these properties, average over systems with fixed values of area per lipids, thickness, and acyl chain order are also shown in the figure. Permeability increases with increasing area per lipid and decreases with increasing membrane thickness and acyl chain order. Previous simulation and experimental studies have reported linear dependence of permeability on both area per lipid and thickness for limited set of lipid mixtures but not always for all types of lipid mixtures \cite{mathai08,frallicciardi22}. Our results suggest stronger dependence for thicknesses below 3.9\,nm and areas above 69\,Å$^2$, but the dependence may appear linear above and below these values as shown in the insets of Figs. \label{fig:permeability} a) and b). As expected, the permeability increases with increasing temperature in Figs. \label{fig:permeability} d). The Arrhenius plot gives 17\,$\pm$\,4\,k$_b$T for the average energy barrier for the water permeation. The values for water permeabilities from simulations in the databank vary between 0.3 and 322\,$\mu$m/s with the mean and median of 14\,$\mu$m/s and 8\,$\mu$m/s, respectively. These values are within the range of experimentally reported values and the mean is the same order of magnitude as the diffusive permeability which is 1-2 orders of magnitude slower than osmotic permeability \cite{jansen95,??}. Clear dependencies of permeability on hydration level, charged lipid fraction, cholesterol fraction or POPE fraction were not observed, although weak decrease for the latter two may be visible.

\begin{figure}[tbp]
    \centering
    \includegraphics[width=180mm]{Figures/permeation2.pdf}
    \caption{Water permeation through lipid membranes analyzed from the databank as a function of 
      a) area per lipid,
      b) thickness,
      and c) temperature. 
      Inserts in a) and b) show the region where the dependence could be considered approximately linear.
      Insert in c) shows the Arrhenius plot of permeation ($\ln$ (P) vs. 1/T) that gives 17\,$\pm$\,4\,k$_b$T for the average activation energy for water permeation through lipid bilayer.}
    \label{fig:permeability}
\end{figure}

To analyze the anisotropic water diffusion, we also calculated the water diffusion along the membrane. The results at different hydration levels are shown in Fig.~\ref{fig:permeability} e) together with the experimental data~\cite{??}. The experimental water diffusion increases with increasing hydration level toward the value for bulk water (?? in ??K). Simulations overestimate the experimental bulk value approximately with the factor of 2 at high hydration levels which is not surprising as the most common water model used in membrane simulations, TIP3P, overestimates the bulk water diffusion. However, simulation results are closer to experiments with low hydration levels. To estimate the diffusion anistoropy, we translated the water permation to diffusion coefficent through multilamellar stack using the Tanner equation~\cite{??}. The results in
